---
title: "NHIS Data"
author: "Kieran Yuen"
font-family: Corbel
output:
  github_document:
    toc: true
---
# Setup
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading in NHIS data
```{r Loading in NHIS data}
library(readxl)
student_datav1 <- read_excel("student_datav1.xlsx")
```

# Cleaning Data
```{r Cleaning Data}
library(tidyverse)
NHIS_data <- student_datav1 %>% 
  select(educ_a,
         hisp_a,
         raceallp_a,
         agep_a,
         sex_a,
         marstat_a,
         phstat_a,
         covidtest_a) %>%
  rename(Education = educ_a,
         Hisp = hisp_a,
         Race = raceallp_a,
         Age = agep_a,
         Sex = sex_a,
         Marital_Status = marstat_a,
         Health_Status = phstat_a,
         Covid_Test= covidtest_a) %>%
  
  filter(Covid_Test =="1"| #Yes_Covid_Test
           Covid_Test == "2") %>% #No_Covid_Test 
  
  mutate(Covid_Test = recode(Covid_Test,
                            '1' = 1,#Yes_Covid_Test
                            '2' = 0)) %>% #No_Covid_Test 
                              
  mutate(Education = recode(Education,
                            '0' = 0,#Below_Bachelors
                            '1' = 0,#Below_Bachelors
                            '2' = 0,#Below_Bachelors
                            '3' = 0,#Below_Bachelors
                            '4' = 0,#Below_Bachelors
                            '5' = 0,#Below_Bachelors
                            '6' = 0,#Below_Bachelors
                            '7' = 0,#Below_Bachelors
                            '8' = 1,#Bachelors_Or_Above
                            '9' = 1,#Bachelors_Or_Above
                            '10' = 1,#Bachelors_Or_Above
                            '11' = 1,#Bachelors_Or_Above
                            '97' = NA_real_,
                            '99' = NA_real_)) %>%
  
  mutate(Hisp = recode(Hisp,
                            '1' = "Hispanic",
                            '2' = "Non_hispanic")) %>% 
  
  mutate(Race = recode(Race,
                            '1' = "White only",
                            '2' = "Black/African American only",
                            '3' = "Asian only",
                            '4' = "AIAN only",
                            '5' = "AIAN and any other group",
                            '6' = "Other single and multiple races",
                            '7' = "Refused",
                            '8' = "Not Ascertained",
                            '9' = "Don't know")) %>%
  
  mutate(Race_Hisp = ifelse(Hisp == 'Non_hispanic' & Race == 'Black/African American only', 
                            'Non-Hispanic Black',
                            
                            ifelse(Hisp == 'Non_hispanic' & Race == 'White only', 'Non-Hispanic White',
                                   
                                   ifelse(Hisp == 'Non_hispanic' & Race == 'AIAN and any other group'|
                                            Hisp == 'Non_hispanic' & Race == 'AIAN only'|
                                            Hisp == 'Non_hispanic' & Race == 'Asian only'|
                                            Hisp == 'Non_hispanic' & Race == 'Other single and multiple races', 
                                          'Other',
                                          
                                          'Hispanic')))) %>%
                            
  mutate(Race_Hisp = recode(Race_Hisp,
                            "Non-Hispanic Black" = 0, #Other
                            "Non-Hispanic White" = 1, #Non_Hispanic_White
                            "Other" = 0,  #Other
                            "Hispanic" = 0)) %>%  #Other
  
  mutate(Sex = recode(Sex,
                            '1' = 1, #Male
                            '2' = 0, #Female
                            '7' = NA_real_,
                            '9' = NA_real_)) %>%
  
  mutate(Marital_Status = recode(Marital_Status,
                            '1' = 1, #Married
                            '2' = 1, #Married
                            '3' = 0, #Not_Married
                            '4' = 0, #Not_Married
                            '5' = 0, #Not_Married
                            '6' = 0, #Not_Married
                            '7' = 0, #Not_Married
                            '8' = 0, #Not_Married
                            '9' = 0)) %>%  #Not_Married
  
  mutate(Health_Status = recode(Health_Status,
                            '1' = 1, #High_Health_Status
                            '2' = 1, #High_Health_Status
                            '3' = 1, #High_Health_Status
                            '4' = 0, #Low_Health_Status
                            '5' = 0, #Low_Health_Status
                            '7' = NA_real_,
                            '9' = NA_real_,
                            )) %>%
  
  mutate(Age_Group = case_when(
    Age <= 34            ~ "27-34",
    Age > 34 & Age <= 44 ~ "35-44",
    Age > 44 & Age <= 54 ~ "45-54",
    Age > 54 & Age <= 64 ~ "55-64"
    ))   %>%
  
  na.omit()
```



# Changing Data Types for variables
```{r Changing Data Types for variables, results="hide"}
NHIS_data$Education <- as.factor(NHIS_data$Education)

NHIS_data$Race_Hisp <- as.factor(NHIS_data$Race_Hisp)

NHIS_data$Age <- as.numeric(NHIS_data$Age)

NHIS_data$Sex <- as.factor(NHIS_data$Sex)

NHIS_data$Marital_Status <- as.factor(NHIS_data$Marital_Status)

NHIS_data$Health_Status <- as.factor(NHIS_data$Health_Status)

NHIS_data$Covid_Test <- as.factor(NHIS_data$Covid_Test)

NHIS_data$Age_Group <- as.factor(NHIS_data$Age_Group)
```


# Summary tables to check groupings are accurate
```{r Summary tables to check groupings are accurate, results="hide"}
table(NHIS_data$Education)
table(NHIS_data$Hisp)
table(NHIS_data$Race)
table(NHIS_data$Race_Hisp)
table(NHIS_data$Sex)
table(NHIS_data$Marital_Status)
table(NHIS_data$Health_Status)
table(NHIS_data$Covid_Test)
table(NHIS_data$Age_Group)
```


# Summary Statistics
## Summary Statistics: COVID Test
```{r Summary Statistics: COVID Test, results="hide"}
library(gmodels)

CrossTable(NHIS_data$Education, NHIS_data$Covid_Test, prop.chisq=FALSE, prop.t=FALSE, prop.r=TRUE, prop.c=FALSE)

CrossTable(NHIS_data$Race_Hisp, NHIS_data$Covid_Test, prop.chisq=FALSE,  prop.t=FALSE, prop.r=TRUE, prop.c=FALSE)

CrossTable(NHIS_data$Sex,NHIS_data$Covid_Test, prop.chisq=FALSE,  prop.t=FALSE, prop.r=TRUE, prop.c=FALSE)

CrossTable(NHIS_data$Marital_Status,NHIS_data$Covid_Test, prop.chisq=FALSE,  prop.t=FALSE, prop.r=TRUE, prop.c=FALSE)

CrossTable(NHIS_data$Health_Status,NHIS_data$Covid_Test, prop.chisq=FALSE,  prop.t=FALSE, prop.r=TRUE, prop.c=FALSE)

CrossTable(NHIS_data$Age_Group,NHIS_data$Covid_Test, prop.chisq=FALSE,  prop.t=FALSE, prop.r=TRUE, prop.c=FALSE)
```

## Summary Statistics: Health Status
```{r Summary Statistics: Health Status, results="hide"}
library(gmodels)

CrossTable(NHIS_data$Covid_Test,NHIS_data$Health_Status, prop.chisq=FALSE,  prop.t=FALSE, prop.r=TRUE, prop.c=FALSE)

CrossTable(NHIS_data$Marital_Status,NHIS_data$Health_Status, prop.chisq=FALSE,  prop.t=FALSE, prop.r=TRUE, prop.c=FALSE)

CrossTable(NHIS_data$Sex,NHIS_data$Health_Status, prop.chisq=FALSE,  prop.t=FALSE, prop.r=TRUE, prop.c=FALSE)

CrossTable(NHIS_data$Age_Group,NHIS_data$Health_Status, prop.chisq=FALSE,  prop.t=FALSE, prop.r=TRUE, prop.c=FALSE)

CrossTable(NHIS_data$Race_Hisp, NHIS_data$Health_Status, prop.chisq=FALSE,  prop.t=FALSE, prop.r=TRUE, prop.c=FALSE)

CrossTable(NHIS_data$Education, NHIS_data$Health_Status, prop.chisq=FALSE, prop.t=FALSE, prop.r=TRUE, prop.c=FALSE)
```


## Summary Statistics: Overall
```{r Summary Statistics: Overall, results="hide"}
library(psych)
describeBy(NHIS_data, group = NHIS_data$Health_Status)
describeBy(NHIS_data)

```






## Summary Statistics: Prof.De's
```{r Summary Statistics: Prof.De, results="hide"}
library(tidyverse)
#Health Status
NHIS_data %>% 
  count(Health_Status) %>% 
  mutate(percent = n / sum(n))

sd(as.numeric(as.factor(NHIS_data$Health_Status==1)))
sd(as.numeric(as.factor(NHIS_data$Health_Status==0)))


#Education
NHIS_data %>% 
  count(Education) %>% 
  mutate(percent = n / sum(n))

sd(as.numeric(as.factor(NHIS_data$Education==1)))
sd(as.numeric(as.factor(NHIS_data$Education==0)))

#Race
NHIS_data %>% 
  count(Race_Hisp) %>% 
  mutate(percent = n / sum(n))

sd(as.numeric(as.factor(NHIS_data$Race_Hisp==1)))
sd(as.numeric(as.factor(NHIS_data$Race_Hisp==0)))

#Sex
NHIS_data %>% 
  count(Sex) %>% 
  mutate(percent = n / sum(n))

sd(as.numeric(as.factor(NHIS_data$Sex==1)))
sd(as.numeric(as.factor(NHIS_data$Sex==0)))

#Marital Status
NHIS_data %>% 
  count(Marital_Status) %>% 
  mutate(percent = n / sum(n))

sd(as.numeric(as.factor(NHIS_data$Marital_Status==1)))
sd(as.numeric(as.factor(NHIS_data$Marital_Status==0)))

#Covid Test
NHIS_data %>% 
  count(Covid_Test) %>% 
  mutate(percent = n / sum(n))

sd(as.numeric(as.factor(NHIS_data$Covid_Test==1)))
sd(as.numeric(as.factor(NHIS_data$Covid_Test==0)))

```









# Logistic Regression
## Model #1: Demographic information only
```{r Model #1: Demographic information only}
Model_1 <- glm(Covid_Test ~ Education
               + Age 
               + Sex 
               + Marital_Status
               + Race_Hisp,
               data = NHIS_data,
               family = binomial)

summary(Model_1)

library(questionr)
odds.ratio(Model_1) #This tells us the "Odds Ratios"

nobs(Model_1)
```



## Model #2: Demographic + Health Status
```{r Model #2: Demographic + Health Status}
Model_2 <- glm(Covid_Test ~ Education
               + Age 
               + Sex 
               + Marital_Status
               + Race_Hisp
               + Health_Status,
               data = NHIS_data,
               family = binomial)

summary(Model_2)

library(questionr)
odds.ratio(Model_2) #This tells us the "Odds Ratios"

nobs(Model_2)
```






## Model #2a: Demographic + Health Status (with only "Other" as Race)
```{r Model #2a: Demographic + Health Status (with only "Other" as Race)}
Model_2a <- glm(Covid_Test ~ Education
               + Age 
               + Sex 
               + Marital_Status
               + Health_Status,
               data = subset(NHIS_data,
                             Race_Hisp == 0),
               family = binomial)

summary(Model_2a)

library(questionr)
odds.ratio(Model_2a) #This tells us the "Odds Ratios"

nobs(Model_2a)
```




## Model #2b: Demographic + Health Status (with only "Non_Hispanic_White" as Race)
```{r Model #2b: Demographic + Health Status (with only "Non_Hispanic_White" as Race)}
Model_2b <- glm(Covid_Test ~ Education
               + Age 
               + Sex 
               + Marital_Status
               + Health_Status,
               data = subset(NHIS_data,
                             Race_Hisp == 1),
               family = binomial)

summary(Model_2b)

library(questionr)
odds.ratio(Model_2b) #This tells us the "Odds Ratios"

nobs(Model_2b)
```
